Transferring Value Functions by Variational Methods

### 2018-05-18 NIPS 2018/main.tex
Changed title

Firstly, we show the results of the evaluation performance. We compute this as the average reward gotten when the agent acts using the greedy policy using the parameters at the iteration of evaluation. In the case of GVT, we take the mean of the posterior Gaussian and, in MGVT, we compute the mean of the posterior by averaging the means of its components and use it for evaluation.

@@

\begin{figure*}[t]
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/three-room/erew.pdf}
    \caption{}
    \label{fig:3r-erew}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/three-room-gen/erew.pdf}
    \caption{}
    \label{fig:3rg-erew}
  \end{subfigure}
  \caption{(a) Rooms Problem: Evaluation Reward, (b) Rooms Problem: Evaluation Reward in the generalization experiment}
  \label{fig:rooms-erew}
\end{figure*}

Furthermore, we investigate the exploratory behavior induced by our transfer algorithms and how they differ between each other and in comparison with $\epsilon$-greedy exploration. In Figure~\ref{fig:rooms-expl}, we show the results of running the no transfer (NT) algorithm, GVT and 1-MGVT for $2k$ iterations and we represent in the plot the positions visited by the agent.

Observing Figure~\ref{fig:nt-vs-gvt}, it is possible to understand the difference between the $\epsilon$-greedy exploration and the resulting behavior from GVT. It is noticeable that NT is not capable to lead the agent to the goal within the given iterations as most of the states visited are sparse within the first room, whereas GVT is able to concentrate more of its effort in looking for the door around the middle of the wall. After finding it, already in the second room, the positions concentrate in the path leading to the goal given that the need for the exploration is less. This is not surprising as the value function should be equal for all tasks after crossing the door.

In the other case, we have Figure~\ref{fig:nt-vs-mgvt} that shows the same situation as with GVT, but it is quite interesting to notice how sparser the exploration of 1-MGVT is with respect to GVT. Indeed, 1-MGVT is able to actually explore the right part of the first room within these iterations, which might be seen as the result of the prior model being able to capture more information than the Gaussian; hence, the higher speed-up in convergence and robustness to changes in the distribution from which target tasks are drawn. Indeed, as 1-MGVT is able to allow for more flexible exploration, it is capable to discover how to best solve the task much faster than GVT.

\begin{figure*}[t]
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/three-room-app/nt_vs_gvt.pdf}
    \caption{}
    \label{fig:nt-vs-gvt}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/three-room-app/nt_vs_mgvt.pdf}
    \caption{}
    \label{fig:nt-vs-mgvt}
  \end{subfigure}
  \caption{(a) Two-Rooms Problem: $\epsilon$-greedy vs. GVT, (b) Two-Rooms Problem: $\epsilon$-greedy vs. 1-MGVT experiment}
  \label{fig:rooms-expl}
\end{figure*}

@@

\begin{figure*}[t]
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/sequential/lrew_1000.pdf}
    \caption{}
    \label{fig:seq-1000}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/sequential/lrew_1950.pdf}
    \caption{}
    \label{fig:seq-1950}
  \end{subfigure}
  \caption{Expected return w.r.t. to the number of source tasks (a) $1000$ iterations, (b) $1950$ iterations}
  \label{fig:sequential}
\end{figure*}

@@

During the training of the source tasks, we use a \textit{Batch Size} of $150$ and \textit{Buffer Size} of $50000$. Specifically, for DDQN we use a \textit{Target Update Frequency} of $500$ and and \textit{Exploration Fraction} of $0.35$ and a learning rate $\alpha=0.001$. We use a Multilayer Perceptron (MLP) with ReLU as activation function and a hidden layer of $32$ neurons, an input layer of $4$ and the value of the $2$ actions as the output layer.

For the transfer experiments, we set the \textit{Batch Size} to $500$, the number of \textit{weights} sampled to approximate the expected TD error to $5$, $\lambda=0.001$ and $\psi=0.5\ $. We use $\alpha_{\mu}=0.001$ as the learning rate for the mean of the Gaussian posterior. For its the Cholesky factor L we use $\alpha_L=0.0001$ and set the limit that the minimum eigenvalue may reach to $\sigma_{min}^2=0.0001\ $. Additionally, for MGVT we set the variance of the prior components $\sigma_p^2=10^{-5}$and leave the learning rates of the posterior components' means and Cholesky factor the same as GVT.

@@

\begin{figure*}[t]
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/cartpole/erew.pdf}
    \caption{}
    \label{fig:cp-erew}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/mountaincar/erew.pdf}
    \caption{}
    \label{fig:mc-erew}
  \end{subfigure}
  \caption{(a) Cartpole: Evaluation Reward, (b) Mountain Car: Evaluation Reward}
  \label{fig:ccerew}
\end{figure*}

@@

For the sources, we train the tasks using DDQN with a \textit{Target Update Frequency} of $500$, a \textit{Batch Size} of $32$, a  \textit{Buffer Size} of $50000$ and learning rate $\alpha=0.001$. Moreover, we set the \textit{Exploration Fraction} to $0.15$. We use an MLP with single hidden layer of $64$ neurons with ReLU activation function, an input layer of $4$ neurons and as output the value of each of the $3$ actions.

For the transfer experiments, we set the \textit{Batch Size} to $500$, and use $10$ \textit{weights} to approximate the expected TD error, $\lambda=10^{-5}$ and $\psi=0.5$. For the learning rates, we use $\alpha_{\mu}=0.001$ for learning the means of the Gaussians. In the case of the Cholesky factors L, we use $\alpha_L=0.0001$ and allow the eigenvalues to reach a minimum value of $\sigma_{min}^2=0.0001$. In the case of MGVT, additionally, we set the prior covariance to be $\sigma_p^2=10^{-5}$.

@@

For the experiments we use as an approximator an MLP with two hidden layers of $32$ neurons with ReLU as activation functions, an input layer receiving the $22$-dimensional state of the MDP and an output layer that produces the value of each of the $3$ action. For training the sources we use DDQN with a \textit{Batch size} of $70$, a \textit{Buffer Size} of $10000$ and a \textit{Target Update Frequency} of $100$, setting the \textit{Exploration Fraction} to $0.1$ and learning rate $\alpha=0.001$ .

@@

% Placer-holder to actual plots. Add analyses
\begin{figure*}[t]
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/maze/lrev.pdf}
    \caption{}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/maze/erew.pdf}
    \caption{}
  \end{subfigure}
  \caption{Maze \ref{fig:maze0} (a) Expected Return (b) Evaluation Reward}
\end{figure*}

\begin{figure*}[t]
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/maze/lrev.pdf}
    \caption{}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/maze/erew.pdf}
    \caption{}
  \end{subfigure}
  \caption{Maze \ref{fig:maze6} (a) Expected Return (b) Evaluation Reward}
\end{figure*}

\begin{figure*}[t]
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/maze/lrev.pdf}
    \caption{}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/maze/erew.pdf}
    \caption{}
  \end{subfigure}
  \caption{Maze \ref{fig:maze13} (a) Expected Return (b) Evaluation Reward}
  \label{fig:mazes-plots}
\end{figure*}

>>>>

Firstly, we show the results of the evaluation of the greedy performance. We compute this as the average reward gotten when the agent acts using the greedy policy using the parameters at the iteration of evaluation. In the case of GVT, we take the mean of the posterior Gaussian and, in MGVT, we compute the mean of the posterior by averaging the means of its components and use it for evaluation.

@@

\begin{figure*}[t]
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.3cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/three-room/erew.pdf}
    \caption{}
    \label{fig:3r-erew}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.3cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/three-room-gen/erew.pdf}
    \caption{}
    \label{fig:3rg-erew}
  \end{subfigure}
  \caption{(a) Rooms Problem: Expected Return w.r.t. Greedy Policy, (b) Rooms Problem: Expected Return w.r.t. Greedy Policy in the generalization experiment}
  \label{fig:rooms-erew}
\end{figure*}

Furthermore, we investigate the exploratory behavior induced by our transfer algorithms and how they differ between each other and in comparison with $\epsilon$-greedy exploration. In Fig. \ref{fig:rooms-expl}, we show the results of running the no transfer (NT) algorithm, GVT and 1-MGVT for $2000$ iterations and we represent as a scatter plot the positions visited by the agent.

Observing Fig. \ref{fig:nt-vs-gvt}, it is possible to understand the difference between the $\epsilon$-greedy exploration and the resulting behavior from GVT. It is noticeable that NT is not capable to lead the agent to the goal within the given iterations as most of the states visited are sparse within the first room, whereas GVT is able to concentrate more of its effort in looking for the door around the middle of the wall. After finding it, within the second room, the positions concentrate in the path leading to the goal given that the need for the exploration is less. This is not surprising as the value function should be equal for all tasks after crossing the door.

In the other case, we have Fig. \ref{fig:nt-vs-mgvt} that shows a similar situation to that of GVT, but it is quite interesting to notice how sparser the exploration of 1-MGVT is with respect to GVT. Indeed, 1-MGVT is able to actually explore the right part of the first room within these iterations, which might be seen as the result of the prior model being able to capture more information than the Gaussian; hence, the higher speed-up in convergence and robustness to changes in the distribution from which target tasks are drawn. Indeed, as 1-MGVT is able to allow for more flexible exploration, it is capable to discover how to best solve the task much faster than GVT.

\begin{figure*}[t]
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0.5cm 1.5cm 1.3cm,clip=true,height=4.5cm]{images/three-room-app/nt_vs_gvt.png}
    \caption{}
    \label{fig:nt-vs-gvt}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.5cm 0.5cm 1.5cm 1.3cm,clip=true,height=4.5cm]{images/three-room-app/nt_vs_mgvt.png}
    \caption{}
    \label{fig:nt-vs-mgvt}
  \end{subfigure}
  \caption{(a) Two-Rooms Problem: $\epsilon$-greedy vs. GVT, (b) Two-Rooms Problem: $\epsilon$-greedy vs. 1-MGVT experiment}
  \label{fig:rooms-expl}
\end{figure*}

@@

\begin{figure*}[t]
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.0cm 0cm 1.2cm 1.3cm,clip=true,height=4.5cm]{images/sequential/lrew_1000.pdf}
    \caption{}
    \label{fig:seq-1000}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.0cm 0cm 1.2cm 1.3cm,clip=true,height=4.5cm]{images/sequential/lrew_1950.pdf}
    \caption{}
    \label{fig:seq-1950}
  \end{subfigure}
  \caption{Expected return w.r.t. to the number of source tasks (a) $1000$ iterations, (b) $1950$ iterations}
  \label{fig:sequential}
\end{figure*}

@@

During the training of the source tasks, we use a \textit{Batch Size} of $150$ and a \textit{Buffer Size} of $50000$. Specifically, for DDQN we use a \textit{Target Update Frequency} of $500$,\textit{Exploration Fraction} of $0.35$ and a learning rate $\alpha=0.001$. We use a Multilayer Perceptron (MLP) with ReLU as activation function and a single hidden layer of $32$ neurons.

For the transfer experiments, we set the \textit{Batch Size} to $500$, the number of \textit{weights} sampled to approximate the expected TD error to $5$, $\lambda=0.001$ and $\psi=0.5\ $. We use $\alpha_{\mu}=0.001$ as the learning rate for the mean of the Gaussian posterior. For its the Cholesky factor L we use $\alpha_L=0.0001$ and set the limit that the minimum eigenvalue may reach to $\sigma_{min}^2=0.0001\ $. Additionally, for MGVT we set the variance of the prior components $\sigma_p^2=10^{-5}$ and leave the learning rates of the posterior components' means and Cholesky factor the same as GVT.

@@

\begin{figure*}[t]
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.0cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/cartpole/erew.pdf}
    \caption{}
    \label{fig:cp-erew}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.0cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/mountaincar/erew.pdf}
    \caption{}
    \label{fig:mc-erew}
  \end{subfigure}
  \caption{(a) Cartpole: Expected Return w.r.t. Greedy Policy, (b) Mountain Car: Expected Return w.r.t. Greedy Policy}
  \label{fig:ccerew}
\end{figure*}

@@

For the sources, we train the tasks using DDQN with a \textit{Target Update Frequency} of $500$, a \textit{Batch Size} of $32$, a  \textit{Buffer Size} of $50000$ and learning rate $\alpha=0.001$. Moreover, we set the \textit{Exploration Fraction} to $0.15$. We use an MLP with single hidden layer of $64$ neurons with ReLU activation function.

For the transfer experiments, we set the \textit{Batch Size} to $500$, and use $10$ \textit{weights} to approximate the expected TD error, $\lambda=10^{-5}$ and $\psi=0.5$. For the learning rates, we use $\alpha_{\mu}=0.001$ for the means of the Gaussians. In the case of the Cholesky factors L, we use $\alpha_L=0.0001$ and allow the eigenvalues to reach a minimum value of $\sigma_{min}^2=0.0001$. In the case of MGVT, additionally, we set the prior covariance to be $\sigma_p^2=10^{-5}$.

@@

For the experiments we use as an approximator an MLP with two hidden layers of $32$ neurons with ReLU as activation functions..For training the sources we use DDQN with a \textit{Batch size} of $70$, a \textit{Buffer Size} of $10000$ and a \textit{Target Update Frequency} of $100$, setting the \textit{Exploration Fraction} to $0.1$ and learning rate to $\alpha=0.001$.

@@

% Placer-holder to actual plots. Add analyses
\begin{figure*}[h]
	\begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.0cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/maze/lrev_maze0.pdf}
    \caption{Expected Return during learning}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.0cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/maze/erew_maze0.pdf}
    \caption{Expected Return w.r.t. greedy policy}
  \end{subfigure}
  \caption{Performance in Maze \ref{fig:maze0}}
  \label{fig:rew-maze0}

  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.0cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/maze/lrev_maze6.pdf}
    \caption{Expected Return during learning}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.0cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/maze/erew_maze6.pdf}
    \caption{Expected Return w.r.t. greedy policy}
  \end{subfigure}
  \caption{Performance in Maze \ref{fig:maze6}}
  \label{fig:rew-maze6}

  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.0cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/maze/lrev_maze13.pdf}
    \caption{Expected Return during learning}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[trim=0.0cm 0cm 1.8cm 1.3cm,clip=true,height=4.5cm]{images/maze/erew_maze13.pdf}
    \caption{Expected Return w.r.t. greedy policy}
  \end{subfigure}
  \caption{Performance in Maze \ref{fig:maze13} }
  \label{fig:rew-maze13}
\end{figure*}

Hereafter, we present additional results of transferring from $5$ source tasks to the mazes shown in Fig. \ref{fig:maze0}, Fig. \ref{fig:maze6} and Fig. \ref{fig:maze13} using both GVT and MGVT. We show both the evaluation performance, i.e. the average performance obtained evaluating the mean parameters of the posterior in a given iteration of the algorithm by following a greedy policy w.r.t. those parameters, and the expected return during the learning process. All the curves result from averaging $20$ independent runs and randomly sampled sources.

In Fig. \ref{fig:maze0}, \ref{fig:maze6}, \ref{fig:maze13}, we can appreciate that in this more complex transfer setting, MGVT is able to provide significant speed-up in a consistent manner in this subset of mazes. It is also noticeable the bad performance obtained with GVT in all cases. The Gaussian prior model clearly fails to capture enough information to transfer in this setting that result in a negative transfer effect.
